---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
---
#Synopsis
In this analysis of fitbit data collected, we were able to look at the patterns in steps taken per day. We found that the average steps taken per day amounted to 9534 with the median at 10400 steps. The most steps taken in a 5-minute interval is 206 steps. We also found a  clear difference between the weekday and weekend activity patterns. On weekdays this person has a lot more activity in the morning (around 9 AM), and is much more sedentary through out the day. On the weekends this person's activity is more uniform, perhaps due to leisure activities of the weekend.

# Loading and preprocessing the data
 Load the data (i.e. read.csv())
``` {r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "./Wearableactivity.zip")
unzip("./Wearabledata")
data <- read.csv("./activity.csv", colClasses = c("integer","Date","numeric"))
```
# Process/transform the data (if necessary) into a format suitable for analysis
```{r}
library(plyr)
library(dplyr)
library(ggplot2)

#This tranformation is only for the last steps: 
data$Weekend <- weekdays(data$date) == "Saturday" | weekdays(data$date) == "Sunday"
data$Weekend <- factor(data$Weekend, levels = c(F, T), labels = c("Weekday", "Weekend"))
activity <- ddply(data, .(interval, Weekend), summarize, steps = mean(steps, na.rm = T))
```
# What is mean total number of steps taken per day?
### Make a histogram of the total number of steps taken each day
``` {r}
stepsdata <- group_by(data,date)
steps <- summarise(stepsdata, total = sum(steps,na.rm=TRUE))
totalnumberofsteps.plot <- ggplot(steps,aes(date,total)) + geom_bar(stat = "identity",color = "white", width = 1, fill = "black") + labs(y="Steps", x="Date", title = "Total Number of Steps Taken Each Day")
totalnumberofsteps.plot
```
## Calculate and report the mean and median total number of steps taken per day
``` {r}
summary(steps$total)
```

# What is the average daily activity pattern?
##Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
``` {r}
steps_rm_na<-filter(data,!is.na(steps))
stepsdata2<-group_by(steps_rm_na,interval)
steps2<-summarise(stepsdata2,average=mean(steps))
averagedailypattern.plot <- ggplot(data=steps2,aes(x=interval,y=average))+geom_line()+ labs(x ="Interval", y="Mean number of steps", title = "Average Steps By Interval")
averagedailypattern.plot
```
##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
``` {r}
steps2[steps2$average == max(steps2$average),]
```

## Imputing missing values
### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(data))
```
## Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
### Create a new dataset that is equal to the original dataset but with the missing data filled in.
``` {r}
data3 <- data 
for (i in 1:nrow(data3)) {
    if (is.na(data3$steps[i])) {
        data3$steps[i] <- steps2[which(data3$interval[i] == steps2$interval), ]$average
    }
}
```
### Display the first few rows of the new activity data frame:
``` {r}
head(data3)
```
# Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
groupSteps3 <- group_by(data3, date)
steps3 <- summarise(groupSteps3,total=sum(steps))
ggplot(steps3,aes(date,total)) + geom_bar(stat="identity",color ="black",fill="black", width = 0.7)+ labs(title = "Histogram of Total Number of Steps Taken Each Day (no missing data)", x = "Date", y = "Total number of steps")
```
### Compute the mean and median.
``` {r}
mean(data3$steps)
median(data3$steps)
```

# Are there differences in activity patterns between weekdays and weekends?
### Create a new factor variable in the dataset with two levels - “weekdays” and “weekend” indicating whether a given date is a weekday or weekend day.
``` {r}
library(lattice)
data$Weekend <- weekdays(data$date) == "Saturday" | weekdays(data$date) == "Sunday"
data$Weekend <- factor(data$Weekend, levels = c(F, T), labels = c("Weekday", "Weekend"))
activity <- ddply(data, .(interval, Weekend), summarize, steps = mean(steps, na.rm = T))
weekday_weekend.plot <- xyplot(steps ~ interval | Weekend, activity, type = "l", layout = c(1, 2), ylab = "Number of Steps", xlab = "Interval", main = "Time Series for Weekend and Weekday Activity Patterns")
weekday_weekend.plot
```